---
title: "Orchestration Load Indicators and Patterns"
subtitle: "Mixed-method Studies Combining Physiological, Subjective and Behavioral Data" 
author: "Luis P. Prieto, Kshitij Sharma & Pierre Dillenbourg"
date: "November 11, 2015"
output: html_document
---

## Executive Summary

Orchestration load (the effort a teacher spends in coordinating the multiple activities and learning processes in an educational setting) has been proposed as a helpful construct in designing educational technologies that not only support learning, but also are usable at the classroom level. Until now, this notion has remained fuzzy, used solely on a high-level and abstract way. In order to ground orchestration load in empirical evidence and study it in a more systematic and detailed manner, we propose a mixed-methods approach combining physiological, behavioral and subjective data, at different time scales (from second-level to multi-session). We also present the results of applying this method to four case studies, where teachers orchestrated technology-enhanced face-to-face lessons with primary school and university students. By looking at the commonalities of our findings across these contexts, we extract evidence-based insights about classroom orchestration with technology, which have potential for generalization (e.g., the increased difficulty of class-level monitoring and interaction). From the differences among the studies we also gain an understanding of the limits of orchestration load as a unidimensional construct, and we elicit factors that seem to influence it most prominently (e.g., teacher experience, the learning design of the lesson).


# A Mixed-method Approach to Study Teacher Orchestration Load

Describe briefly the method...


```{r, child="study1alt.Rmd"}
```


```{r, child="study2alt.Rmd"}
```

```{r, child="study3alt.Rmd"}
```


# Conclusion

Outline briefly some of the main discussion, interpretation... (then put nice prose in the paper)