# Study 1: Feasibility of the Approach (Open Doors Day)

In this case study, we aimed to explore the following research questions: _is it feasible to use a mobile eye-tracker to follow a teacher's cognitive load in a semi-authentic classroom setting? Does such analysis provide interesting insights about classroom usability of a novel technology? Can we detect specific classroom interaction episodes that imply high (or low) cognitive load?_.

The **context** of the data gathering were three sessions of 35-45 minutes, in a classroom with several tabletop computers, in which classes of around 20 students were playing collaboratively to a tangible tabletop game, to learn about fractions, fraction comparison and equivalence. The teacher was wearing a mobile eye-tracker

## Before starting: Data download and pre-process

First of all, we download the datasets for the three studies, which have been published in Zenodo: 

* [Dataset for Study 1 (Multi-tabletop sessions in the lab, primary school students)](https://zenodo.org/record/16515)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
# We load the useful scripts and packages needed throughout the report
source("./lib/rollingWindows.R")
source("./lib/loadIndex.R")
source("./lib/extremeLoadExtraction.R")
source("./lib/aggregateEpisodeData.R")
source("./lib/multiplot.R")
source("./lib/outliers.R")

rootdir <- getwd()

# If not present already, download dataset study 2 and uncompress it to Study2/
setwd(paste(rootdir,"/data/study1",sep=""))
if(!file.exists("JDC2014-VideoCodingData.zip") || !file.exists("JDC2014-EyetrackingData.zip")){
    download.file("https://zenodo.org/record/16515/files/JDC2014-VideoCodingData.zip", destfile="JDC2014-VideoCodingData.zip", method="curl")
    unzip("JDC2014-VideoCodingData.zip")
    download.file("https://zenodo.org/record/16515/files/JDC2014-EyetrackingData.zip", destfile="JDC2014-EyetrackingData.zip", method="curl")
    unzip("JDC2014-EyetrackingData.zip")
} 

# Now we have the raw data files uncompressed in the data/study1 folder
```

Once we have the raw data from the eye-tracker (plus the video coding data generated by researchers), we pre-process it, mostly by aggregating the four load-related eyetracking metrics (pupil diameter mean, pupil diameter variation, saccade speed and number of fixations >500ms) into 10-second episodes, using a rolling window with 5-second slide between windows (see ```./lib/aggregateEpisodeData.R``` and ```./lib/rollingWindows.R``` files for details).

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
setwd(paste(rootdir,"/data/study1",sep=""))

# We do the preprocessing, which will generate a Rda file with the 10s
# window data, and will return the name of the file

sessions <- c("JDC2014-Session1-eyetracking","JDC2014-Session2-eyetracking","JDC2014-Session3-eyetracking")

cleandatafile <- "study1ProcessedData.Rda"
if(!file.exists(cleandatafile)){
    data <- aggregateEpisodeData(sessions, datadir=paste(rootdir,"/data/study1",sep=""), initendtimes=NULL, SEPARATOR=",") # For this study the raw data is comma-separated!
    data <- data[,c(1:5,12)] # We select only the load-related metrics
    # We load and add the video coding data with the social, activity and main gaze focus dimensions
    videocodes <- data.frame()
    sessionsvid <- c("JDC2014-Session1","JDC2014-Session2","JDC2014-Session3") # For some reason, the filenames for the videocoding are not consistent with the previous session labels
    for(session in sessionsvid){
        sessioncodes <- read.csv(paste(session,"-videocoding.csv",sep = ""), sep=",")
        if(nrow(videocodes)==0) videocodes <- sessioncodes
        else videocodes <- rbind(videocodes,sessioncodes)
    }
    videocodes$session <- videocodes$Session
    totaldata <- merge(data,videocodes,by=c("session","time"),all=T)
    save(totaldata, file=cleandatafile)
}

```

### A: From samples to episodes

We first calculate the _coarse load index_ (how many measures are over the median) and the _fine load index_ (average of percentiles of the different load metrics), and we plot them for each session

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3, fig.width=8}
setwd(paste(rootdir,"/data/study1",sep=""))
totaldata <- get(load(cleandatafile))

# We crop outliers in the saccade speed. They tend to throw off the HMM and PCA calculations
#countOutliers(totaldata$value.Sac,coef = 5, method="iqr") # 5xinter-quartile range
totaldata$value.Sac <- replaceOutliers(totaldata$value.Sac,valueNA = F,coef = 5, method="iqr")

loaddata <- calculateCoarseFineLoadIndex(totaldata,3:6,normalize=T) # We ensure that the values are normalized for those of the first window in which little load is assumed (just putting the eyetracker on)

# TODO: Check that all load metrics are directly proportional. If not, we need to introduce a modification in the calculations of load

# We plot the loads for each session, along with some smoothing
for(session in sessions){
    sessiondata <- loaddata[loaddata$session==session,]
    p1 <- ggplot(sessiondata, aes(x=time, y=CoarseLoad, col=CoarseLoad)) + 
            ggtitle(paste("Coarse Load Index ",session,sep="")) + 
            geom_line(size=1) + stat_smooth(method="loess",span=0.1,se=F) +
            #theme(axis.text.x = element_text(size=18),plot.title=element_text(size=20, face="bold"),axis.title=element_text(size=18),panel.background = element_rect(fill = 'white')) +
            theme(panel.background = element_rect(fill = 'white')) +
            scale_color_gradient(low="green",high="red")
    print(p1)

    p2 <- ggplot(sessiondata, aes(x=time, y=FineLoad, col=FineLoad)) + 
            ggtitle(paste("Fine Load Index ",session,sep="")) + 
            geom_line(size=1) + stat_smooth(method="loess",span=0.1,se=F) +
            #theme(axis.text.x = element_text(size=18),plot.title=element_text(size=20, face="bold"),axis.title=element_text(size=18),panel.background = element_rect(fill = 'white')) +
            theme(panel.background = element_rect(fill = 'white')) +
            scale_color_gradient(low="green",high="red")
    print(p2)
}
```

### B: Exploring the structure of the sensor (episode) load data 

First we do a Principal Component analysis to see what dimensions are there within these 4 load-related signals:

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5, fig.width=5}

library(FactoMineR)

# Overall for the three sessions
res.pca = PCA(loaddata[, c(3:6)], scale.unit=TRUE, ncp=5, graph=F)
plot.PCA(res.pca, axes=c(1, 2), choix="var", title="Overall data")
loaddata$PupilLoadDim = res.pca$ind$coord[,1]
loaddata$EyeMoveFixLoadDim = res.pca$ind$coord[,2]


# For each session, results are quite similar
#for(session in sessions){
#    sessiondata <- loaddata[loaddata$session==session,]
#    res.pca = PCA(sessiondata[, c(3:6)], scale.unit=TRUE, ncp=5, graph=F)
#    print(plot.PCA(res.pca, axes=c(1, 2), choix="var",title = paste("Session",session)))
#}
```

We can see all the sessions have pretty similar dimensions, with one dimension taking the pupil diameter-related values, and the other one taking the fixations vs. saccades (in opposing directions, maybe indicating two different ways in which a teacher can be "loaded").


To explore the temporal structure of these load-related sensor signals, we train a Hidden Markov Model with 4 states using the load-related signals as observables:

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5, fig.width=5}
library(depmixS4)

# We plot the BIC for different number of states, just to check
states <- 2:16
BICs <- 0
for(i in states){
#    print(paste(numstates,"..."))
    # Overall
    set.seed(1)
    #set up the HMM model
    mod <- depmix(list(value.Mean~1,value.SD~1,value.Sac~1,value.Fix~1), data=loaddata, nstates=i, family=list(gaussian(),gaussian(),gaussian(),poisson()),ntimes=summary(as.factor(loaddata$session)), verbose=F)
    #fit the model 
    f <- fit(mod, verbose=F)
#    print(f)
    BICs[i] <- BIC(f)
}
qplot(1:16,BICs,geom="line")



# We remove the missing data, as the HMM library does not like them, apparently
# Not needed, we have no missing data in that sense
#data4hmm <- loaddata[complete.cases(loaddata[,c("value.Mean","value.SD","value.Sac","value.Fix")]),]

numstates <- 4

set.seed(1)
#set up the HMM model
mod <- depmix(list(value.Mean~1,value.SD~1,value.Sac~1,value.Fix~1), data=loaddata, nstates=numstates, family=list(gaussian(),gaussian(),gaussian(),poisson()),ntimes=summary(as.factor(loaddata$session)), verbose=F)
#fit the model 
f <- fit(mod, verbose=F)
#check to see how the state transtion matricies and process mean/sd matches our sample data
summary(f)
#get the estimated state for each timestep 
esttrans <- posterior(f)
loaddata$MMstate <- esttrans[,1]
#plot(1:nrow(loaddata), esttrans[,1], type='p', col=as.factor(loaddata$session), main=paste('Estimated state',numstates))

loaddata$MMstate <- factor(loaddata$MMstate)
xtabs(~CoarseLoad+MMstate,data=loaddata)
g1 <- ggplot(data = loaddata, mapping = aes(MMstate, FineLoad))+geom_boxplot(aes(fill=MMstate))+geom_jitter()
print(g1)

# We can also plot the states with regard to the two load dimensions from the PCA
g2 <- ggplot(data = loaddata, mapping = aes(PupilLoadDim,EyeMoveFixLoadDim)) + geom_point(aes(col=MMstate)) + xlim(-4,4) + ylim(-5,5)
print(g2)

#State 1/2 seems to be mid/hi load, state 3 is mid load, state 4 is low-load
levels(loaddata$MMstate) <- c("HiFixLoadState","HiSacLoadState","MidLoadState","LowLoadState")

```

As we can see, States 1 and 2 seem to correspond to high load episodes, state 3 is more mid-load, while state 4 is rather low-load. Also, we see how these states largely serve to separate high-load values based on the pupil diameter metrics, while the 1/2 states seem to capture the variations in the saccadic/fixation dimension of the high-load episodes.

### C: Subjective episode analysis

This part was not done in this study, as the researcher was the main teacher, and might have had assumptions about what each episode's load should be (thus, his subjective data might have been biased)

### D: Behavioral episode analysis

From the coarse load index, we extracted the extreme load episodes (10-second episodes with a coarse load index of 0 or 4), for manual coding of the subjective video feed by a researcher (see the ```./lib/extremeLoadExtraction.R``` file).

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
outputfile <- paste(rootdir,"/data/study1/TimesToVideoCode.csv",sep="")
extremedata <- extractExtremeLoadMoments(sessions, loaddata, outputfile)
```

We manually coded the videos of all these extreme load episodes, along the dimensions of teacher activity, social plane of the interaction and main focus of gaze (see the ```JDC2014-Session*-videocoding.csv``` datafiles). We then aggregate the code counts for each of these dimensions, to bridge from the episode-level analysis to have an idea of what kind of episodes are often high or low (coarse) load. Also, comparisons with fine load, and the PCA and HMM dimensions of load data are made.

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# Overall behavioral analyses
# ACTIVITY
# Additional non-extreme episodes were also coded, we eliminate them to get the coarse load contrast
behdata <- loaddata[(loaddata$CoarseLoad==4|loaddata$CoarseLoad==0)&(!is.na(loaddata$Activity)),]
behdata$Activity <- factor(behdata$Activity)
# chi-squared test and counts/residuals table for teacher activity
tabAct <- table(behdata$CoarseLoad,behdata$Activity)
chisq.test(tabAct)
tabAct
chisq.test(tabAct)$residuals

# Plotting against the FineLoad is not very revealing, as we took extreme values and the averages will always be around the average values!
#g1 <- ggplot(loaddata[!is.na(loaddata$Activity),],aes(Activity,FineLoad)) + geom_boxplot(aes(fill=Activity))
#print(g1)
# Density plots for the PCA dimensions for each kind of activity
# TDT and EXP have higher values in the pupil-diameter load, MON and REP tend to have lower values
g1 <- ggplot(loaddata[!is.na(loaddata$Activity),],aes(PupilLoadDim)) + geom_density(aes(col=Activity))
print(g1)
# EXP tends to have higher values in the fixation kind of dimension, MON and REP have lower values
g2 <- ggplot(loaddata[!is.na(loaddata$Activity),],aes(EyeMoveFixLoadDim)) + geom_density(aes(col=Activity))
print(g2)
# Another way of seeing this, in the PCA diagram
res.pca = PCA(loaddata[, c(3:6, 10)], scale.unit=TRUE, ncp=5, graph=F, quali.sup=5)
#plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5)
plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5, invisible="ind")

# Comparison against the HMM states of load
hmmActTab <- xtabs(~Activity+MMstate, data=loaddata[!is.na(loaddata$Activity),])
chisq.test(hmmActTab)
hmmActTab
chisq.test(hmmActTab)$residuals

```

The **Activity** distribution is unlikely to be independent of the coarse load, but residuals are not very high, they are rather trends not significant by themselves.

The two dimensions serve to see again how TDT and EXP have higher values in the pupil-diameter load, MON and REP tend to have lower values; EXP tends to have higher values in the fixation kind of dimension, MON and REP have lower values.

The HMM states modelling of load help us confirm again that EXP, TDT are high-load states, but also that MON has a trend of high-saccade load episodes. We also see that MON, REP tend to be in the low-load state


```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# SOCIAL
behdata <- loaddata[(loaddata$CoarseLoad==4|loaddata$CoarseLoad==0)&(!is.na(loaddata$Social)),]
behdata$Social <- factor(behdata$Social)
# chi-squared test and counts/residuals table for social plane of interaction
tabSoc <- table(behdata$CoarseLoad,behdata$Social)
chisq.test(tabSoc)
tabSoc
chisq.test(tabSoc)$residuals

# Plotting against the FineLoad is not very revealing, as we took extreme values and the averages will always be around the average values!
g1 <- ggplot(loaddata[!is.na(loaddata$Social),],aes(Social,FineLoad)) + geom_boxplot(aes(fill=Social))
print(g1)
# Density plots for the PCA dimensions for each kind of activity
g1 <- ggplot(loaddata[!is.na(loaddata$Social),],aes(PupilLoadDim)) + geom_density(aes(col=Social))
print(g1)
g2 <- ggplot(loaddata[!is.na(loaddata$Social),],aes(EyeMoveFixLoadDim)) + geom_density(aes(col=Social))
print(g2)
# Another way of seeing this, in the PCA diagram
res.pca = PCA(loaddata[, c(3:6, 11)], scale.unit=TRUE, ncp=5, graph=F, quali.sup=5)
#plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5)
plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5, invisible="ind")

# Comparison against the HMM states of load
hmmSocTab <- xtabs(~Social+MMstate, data=loaddata[!is.na(loaddata$Social),])
chisq.test(hmmSocTab)
hmmSocTab
chisq.test(hmmSocTab)$residuals

```

Again, the distribution with respect to **Social plane** is statistically significant, and the high-load episodes tend to be in the classroom plane, while the low-load tend to be in the group plane. 

The distribution in the two main load dimensions is also very clear, pointing to lower values in the group plane and higher values in the classroom plane. This also means that group-level activities tend to have more saccade-based load (maybe related to MON episodes, see above), while classroom level episodes tend to have more long-fixation-heavy load.

These conclusions are also confirmed by the HMM load model states: classroom level is high load (especially, long-fixation ones), while group-level is more low-mid load states.

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# FOCUS
behdata <- loaddata[(loaddata$CoarseLoad==4|loaddata$CoarseLoad==0)&(!is.na(loaddata$Focus)),]
behdata$Focus <- factor(behdata$Focus)
# chi-squared test and counts/residuals table for main focus of gaze
tabFoc <- table(behdata$CoarseLoad,behdata$Focus)
chisq.test(tabFoc)
tabFoc
chisq.test(tabFoc)$residuals

# Plotting against the FineLoad is not very revealing, as we took extreme values and the averages will always be around the average values!
g1 <- ggplot(loaddata[!is.na(loaddata$Focus),],aes(Focus,FineLoad)) + geom_boxplot(aes(fill=Focus))
print(g1)
# Density plots for the PCA dimensions for each kind of activity ... not very readable
g1 <- ggplot(loaddata[!is.na(loaddata$Focus),],aes(PupilLoadDim)) + geom_density(aes(col=Focus))
print(g1)
g2 <- ggplot(loaddata[!is.na(loaddata$Focus),],aes(EyeMoveFixLoadDim)) + geom_density(aes(col=Focus))
print(g2)
# Another way of seeing this, in the PCA diagram
res.pca = PCA(loaddata[, c(3:6, 12)], scale.unit=TRUE, ncp=5, graph=F, quali.sup=5)
#plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5)
plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5, invisible="ind")

# Comparison against the HMM states of load
hmmFocTab <- xtabs(~Focus+MMstate, data=loaddata[!is.na(loaddata$Focus),])
chisq.test(hmmFocTab)
hmmFocTab
chisq.test(hmmFocTab)$residuals
```

Again, the distribution of **Gaze focus** is very unlikely to be independent of the cognitive load measures. The residuals tell us that high-load episodes are more often looking at backs or faces or people, other experimenters or the teacher table with all the tangibles. Low-load are majoritarily focusing on the tabletop interface itself. 

In this case, both the PCA dimensions analysis and the HMM states analysis clearly support this view of the influence of main gaze focus on load metrics.

Additional, per-session behavioral analyses can be made too (but probably not very enlightening)

### E: Session-level analysis and comparison

Since we had no subjective ratings (for the session or for individual episodes) for cognitive load, all we can do is compare the quantitative, sensor-based load values for each session, to get an idea of which session could have been more or less "loading"

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# Plotting the FineLoad of each session
g1 <- ggplot(loaddata,aes(session,FineLoad)) + geom_boxplot(aes(fill=session))
print(g1)
g1 <- ggplot(loaddata,aes(FineLoad)) + geom_density(aes(col=session))
print(g1)

# Plotting against the PCA and HMM dimensions
g1 <- ggplot(loaddata,aes(PupilLoadDim)) + geom_density(aes(col=session))
print(g1)
g1 <- ggplot(loaddata,aes(EyeMoveFixLoadDim)) + geom_density(aes(col=session))
print(g1)
loaddata$session <- factor(loaddata$session)
res.pca = PCA(loaddata[, c(3:6, 1)], scale.unit=TRUE, ncp=5, graph=F, quali.sup=5)
#plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5)
plot.PCA(res.pca, axes=c(1, 2), choix="ind", habillage=5, invisible="ind")

# Comparison against the HMM states of load
hmmSessTab <- xtabs(~session+MMstate, data=loaddata)
chisq.test(hmmSessTab)
hmmSessTab
chisq.test(hmmSessTab)$residuals



```

We see that, despite differences not being very large, session 2 seemed to have more abundance of high-load episodes.

The PCA dimensions help us see that session 1 was more prone to long-fixation-heavy episodes than the other two. Also, we see how actually session 3 was a little more load from a pupil-diameter dimension

The analysis from the point of view of the HMM states is also interesting, as there are clear differences among otherwise quite similar sessions: session 1 was rich in long-fixation high-load episodes, while sessions 2 and 3 had more saccade-speed based high-load episodes. While session 3 was "average" in mid- and low-load episodes, session 1 had more low-load episodes, and session 2 had more mid-load episodes.

```{r, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# We come to the root folder, for the next study
setwd(rootdir)
```
